---
title: 'Seminoff EPac green turtle Stable Isotope Data Analysis'
author: "Lisa Komoroske"
date: '`r Sys.Date()`'
output: word_document
session info: '`sessionInfo(package = NULL)`'  
---
### Metadata from Jeff  
**Site** - an ordinal code for each site  
**Site code** - 3 letter code for each site  
**Ordered_SITE** - combined site code with ordered # roughly North to South for graphing ordering  
**Location_Label** - shortened location names for graphing labeling purposes  
**Location** - location of turtle capture    
**LAB ID** - self explanatory  
**Collection Date** - self explanatory  
**Run Date** - self explanatory  
**%N** - elemental concentration of N.  that is, how much each sample is made up of nitrogen.  this is used as a diagnostic to know sample quality (anything outside of ~9-17% N raises a red flag)  
**%C** - elemental concentration of C.  that is, how much each sample is made up of carbon.  this is used as a diagnostic to know sample quality (anything outside of ~40-60% C raises a red flag)  
**d15N** - stable isotope value for N  
**d13C** - stable isotope value for C  
**Color** - rarely filled in. This is largely for the Galapagos and Colombia, where black turtles (eastern Pacific stock) and yellow turtles (west Pacific origins) co-exist.  Safe to say that anything that is not filled in here would be a ‘black’ morph.  
**SCL** - straight carapace length  
**CCL_calc_fromSCL** - used formula from Seminoff et al. 2003 to interpolate CCLs from SCLs  
**CCL_empirical** - curved carapace length-these are only the empirically collected values  
**CCL_combined** - curved carapace length-I copied over all the empirical values, and then for ones that were missing empirical CCL but had **CCL_calc_fromSCL**, I added these in; so this is the combined variable that we'll use for size relationships  

### Setup
```{r setup, echo=FALSE,include=FALSE,message = FALSE}
library(here)
setwd(here("scripts&RMD_files"))
getwd()
```

### Load Required Libraries
```{r libraries, echo=FALSE, include=FALSE}
library(ggplot2)
library(plyr)
library(RColorBrewer)#can also use palettes, etc. l=defines darkness/lightness of colors see R graphics cookbook pgs255-263
#additional code and plots from Latimer class for data QC and exploration:####
library(MASS); library(lattice)
library(lme4); library(lmerTest);library(car)
```

### Read in data
```{r read_data, echo=FALSE,include=FALSE}
data<-read.csv("../data/CM Isotope Data UPDATED_Jan122017.csv")
#updated 12-20-16 summary files reading in here to reflect new groupings, excluded data etc.
data$Ordered_SITE<-factor(data$Ordered_SITE, levels =c("1-SGR_SBN","3-SDB","4-NGU","6-BMA","7-IPD","8-LOR","9-BLA","11-CIN","12-NAV", "13-DUL","14-PAR","15-MEJ","16-ISL","17-COC","18-GOR","19-IGP","20-IGE","21-IGD","22-IGZ","23-IGN","24-PPE"))

#12-19-16 added: removing problematic sites and questionable values
data_new<-subset(data, Ordered_SITE!="7-IPD" &Ordered_SITE!="8-LOR"&Ordered_SITE!="16-ISL"&Ordered_SITE!="22-IGZ"&Ordered_SITE!="23-IGN")
data_new<-subset(data_new, Percent_N>=5 | is.na(data_new$Percent_N) )
data_new<-subset(data_new, Percent_N<=20 | is.na(data_new$Percent_N) )
data_new$Collect_Date <- strptime(data_new$Collect_Date, "%m/%d/%Y")
data_new$Run_Date <- strptime(data_new$Run_Date, "%m/%d/%Y")
palette<-c("firebrick","forestgreen","deepskyblue3","purple","orange")
data_new2<-subset(data_new,Habitat_Type!="oceanic") #exclude oceanic site
data<-data_new

data_Nsum<-read.csv("../data/EPGT_d15Nsumc_LKedit.csv")
str(data_Nsum) 
data_Csum<-read.csv("../data/EPGT_d13Csumc_LKedit.csv")
str(data_Csum)
```

### Coarse data QC checks to note obvious data structure problems, etc.:
```{r initial_QC, echo=TRUE,include=FALSE}
#see V1 data exploration scripts for more in depth QC checks
summary(data)
data$Ordered_SITE<-factor(data$Ordered_SITE)#reset variable to get rid of excluded sites in count
table(data$Ordered_SITE) #frequency table
```

### Figures
#### Figure 1. Map  
```{r map, echo=FALSE,include=FALSE, message=FALSE}
# Load map packages
library("maps")
library("mapproj")
library("mapplots")
library("mapdata")
library("ggmap")
#don't need all these packages just for this one map, but I use them later for other map
#stuff in the script that I am not including here

EPGT_latlong<-read.csv("../data/Lat_long_EPGT.csv")
EPGT_latlong$mean_d15N_scale<-EPGT_latlong$mean_d15N*0.4
EPGT_latlong$mean_d15N_scale_1<-EPGT_latlong$mean_d15N*0.75
#Entire Study Region Map
description<-"EPacRegion"
myLocation <- c(-175, -45, -62, 45)
#bounding box lowerleftlon, lowerleftlat, upperrightlon, upperrightlat
myMap <- get_map(location=myLocation, source="google",
                  maptype="hybrid", crop=TRUE) 
myMap
# pdf(paste(description, "_Gmap", ".pdf",sep=""), width = 11, height = 8.5)
# ggmap(myMap) +
#   geom_point(aes(x = Longitude, y = Latitude), 
#              data = EPGT_latlong, fill = "red", alpha=0.65, 
#              size = 6,shape=21)
# dev.off()

#SoCal_Baja_inset
description1<-"SoCal-Baja_region"
SoCal_Baja <- c(-121, 22.5, -109, 35)
#bounding box lowerleftlon, lowerleftlat, upperrightlon, upperrightlat

myMap1 <- get_map(location=SoCal_Baja, 
                  source="google", maptype="satellite", crop=TRUE) 
myMap1
# pdf(paste(description1, "_Gmap", ".pdf",sep=""), width = 8, height = 8.5)
# ggmap(myMap1) +
#   geom_point(aes(x = Longitude, y = Latitude), 
#              data = EPGT_latlong,fill = "red", alpha=0.65, 
#              size = 16, shape=21)
# dev.off()

#Cen-Islands_inset
description2<-"Islands_region"
Islands_region <- c(-100, -10, -75, 11)
#bounding box lowerleftlon, lowerleftlat, upperrightlon, upperrightlat

myMap2 <- get_map(location=Islands_region, 
                  source="google", maptype="satellite", crop=TRUE) 
myMap2
# pdf(paste(description2, "_Gmap", ".pdf",sep=""), width = 8, height = 8.5)
# ggmap(myMap2) +
#   geom_point(aes(x = Longitude, y = Latitude), 
#              data = EPGT_latlong, fill = "red", alpha=0.65,  
#              size = 16, shape=21)
# dev.off()

#So America Region Map
description3<-"SoAm_Region"
SoAm_Region <- c(-81, -25, -68,-13)
#bounding box lowerleftlon, lowerleftlat, upperrightlon, upperrightlat

myMap3 <- get_map(location=SoAm_Region, 
                  source="google", maptype="satellite", crop=TRUE) 
myMap3
# pdf(paste(description3, "_Gmap", ".pdf",sep=""), width = 8, height = 8.5)
# ggmap(myMap3) +
#   geom_point(aes(x = Longitude, y = Latitude), 
#              data = EPGT_latlong, fill = "red", alpha=0.65, 
#              size = 16,shape=21)
# dev.off()
#final figure result of post-processing in inkscape to overlay insets
```

#### Figure 2. Whiskers  
```{r CvsN_whisker, echo=FALSE, message=FALSE}

CNsum_merge<-merge(data_Csum,data_Nsum)
CNsum_merge$SITENS<-paste(CNsum_merge$N_to_S_ordination, "-",CNsum_merge$Ordered_SITE)
CNsum_merge$SITENS
CNsum_merge$SITENS<-factor(CNsum_merge$SITENS,levels= c("1 - MEJ","2 - PAR","3 - PPE" ,"6 - IGD",
"7 - IGE","8 - IGP","9 - GOR", "10 - COC", "11 - DUL","12-NAV", "13 - BMA", "16 - NGU" ,"18 - CIN","19 - BLA","20 - SDB","21 - SGR_SBN")) 

CNsum_merge$Region<-factor(CNsum_merge$Region,levels=c("SC-BC Pac Coast","Gulf of Cal","Cen-SoAm Pac Coast" , "Oceanic" , "EPac Islands"))

reg_palette<-c("forest green","red", "purple","deepskyblue3","orange")
g<-ggplot(CNsum_merge, aes(x=mean_d13C,y=mean_d15N,fill=Region)) +
  geom_errorbar(aes(ymin=mean_d15N-se_d15N,ymax=mean_d15N+se_d15N),size=0.25)+
  geom_errorbarh(aes(xmin=mean_d13C-se_d13C,xmax=mean_d13C+se_d13C),size=0.25)+
  geom_point(size=3, alpha=.9,shape=21)+theme_bw()+theme(axis.text.x  = element_text(angle=0, vjust=0.5))+scale_fill_manual(values=reg_palette)+ylab("Mean d15N per Location")+xlab("Mean d13C per Location")+geom_text(aes(label=Ordered_SITE),hjust=1.5, vjust=0.5,size=3)
g

# pdf("../Output_figures/C vs. N Whisker plot.pdf", 8, 5)
# g
# dev.off()
```

#### Figure 3. See seperate script for Bayesian ellipses and convex hulls
#### Figure 4. CSIA graphic
```{r CSIA, echo=FALSE, message=FALSE }
library(tidyr)
library(dplyr)
CSIA<-read.csv("../data/CSIA Raw data.csv")
CSIA.long<-gather(CSIA,"amino.acid","nitrogen", 3:16) 

N_mean<-tapply(CSIA.long$nitrogen, list(CSIA.long$amino.acid, CSIA.long$Location),mean,na.rm=TRUE)

N_sd<-tapply(CSIA.long$nitrogen, list(CSIA.long$amino.acid, CSIA.long$Location),sd,na.rm=TRUE)

meanN<-data.frame(N_mean)
df <- add_rownames(meanN, "amino.acid")#note this is deprecated in newer versions; update to `tibble::rownames_to_column()` instead
mean.long<-gather(df,"location","mean_N", 2:8) 

sdN<-data.frame(N_sd)
df2 <- add_rownames(sdN, "amino.acid")
sd.long<-gather(df2,"location","sd_N", 2:8) 
summary<-cbind(mean.long,sd.long)
summary<-summary[,c(1:3,6)]

shape<-c(21,22,23,24,25,24,21)
fill<-c("deepskyblue1","mediumblue","red","forestgreen","purple","yellow","grey70")
str(summary)
summary$amino.acid<-factor(summary$amino.acid)
levels(summary$amino.acid)
summary$amino.acid<-factor(summary$amino.acid,levels =c("Ala", "Asx","Glx", "Iso", "Leu", "Pro", "Val", "Gly",   "Lys", "Met", "Phe", "Ser", "Thr" ,"Tyr" ))

summary$location<-factor(summary$location)
levels(summary$location)
summary$location<-factor(summary$location,levels =c("SDB","BLA" ,"BMA", "DUL", "IGD","PAR", "MEJ"))
pd <- position_dodge(1)
g1<-ggplot(data = summary, aes(x = amino.acid, y = mean_N)) + geom_point(size=4, alpha=0.8,aes(shape=location,fill=location)) + 
  labs(x = "Amino Acid", y = "d15N") + theme_classic()+geom_errorbar(aes(ymin=mean_N-sd_N, ymax=mean_N+sd_N), width=.1)
g2<-g1+scale_shape_manual(values=shape,name="Study Site")+scale_fill_manual(values=fill,name="Study Site")+geom_vline(xintercept = 7.5, linetype = "longdash",color="grey30")
g2
# pdf("../Output_figures/CSIA_dotplot1.pdf", 8, 5)
# g2
# dev.off()
```

#### Supplemental Figure 1. Faceted Histograms
```{r histogram_fig, echo=FALSE, message=FALSE}
levels(data$Location_Label)
#come back and remake with codes in parentheses(?)
data$Location_Label<-factor(data$Location_Label, levels =c(
"Long Beach, USA",               "Bahia de los Angeles, GoC, MX",   "Cocos Island, Costa Rica" ,   "Punta Espinosa, Fernandina, Galapagos" , 
"San Diego Bay, USA",            "Infiernillo Channel, GoC, MX" , "Isla Gorgona, Colombia" ,      "Oceanic Waters, Peru (Longline)",  
"North Gulf of Ulloa, BCS, MX",   "Navachiste, GoC, MX",  "Bahia Elizabeth, Isabela, Galapagos",         "Pisco / Paracas Bay, Peru" ,  
"Bahia Magdalena, BCS, MX",       "Golfo Dulce, Costa Rica",  "Caleta Derek, Isabela, Galapagos",    "Mejillones, Chile"        ))
#cocos - Gorgina - Bahia Elizabeth - caleta Derek.  For last column let's go punta e - oceanic Peru - pisco - chile
C<-ggplot(data, aes(x=d13C)) + geom_histogram(alpha=.4, fill="yellow",colour="black")+theme_bw()+
geom_vline(aes(xintercept=mean(d13C, na.rm=T)),    #Ignore NA values for mean
           color="red", linetype="dashed", size=0.5)
C1<-C+facet_wrap( ~ Location_Label, ncol=4)
C1
# pdf("../Output_figures/carbon_faceted_histogram.pdf", 12, 5)
# C1
# dev.off()

N<-ggplot(data, aes(x=d15N)) + geom_histogram(alpha=.4, fill="forest green",colour="black")+theme_bw()+
geom_vline(aes(xintercept=mean(d15N, na.rm=T)),    #Ignore NA values for mean
           color="red", linetype="dashed", size=0.5)
N1<-N+facet_wrap( ~ Location_Label, ncol=4)
N1
# pdf("../Output_figures/nitrogen_faceted_histogram.pdf", 12, 5)
# N1
# dev.off()
```

#### Supplemental Figure 2. Variance Multiplication Factors  
*[come back and insert code or redirect to script]*  
#### Supplemental Figure 3. Boxplots, 2 panel (C then N)  
```{r boxplots, echo=FALSE, message=FALSE}
data$Region<-factor(data$Region,levels=c("SC-BC Pac Coast","Gulf of Cal","Cen-SoAm Pac Coast" , "Oceanic" , "EPac Islands"))
levels(data$Ordered_SITE)
#for bottom to top order
data$Ordered_SITE<-factor(data$Ordered_SITE,levels=c("1-SGR_SBN" ,"3-SDB",     "4-NGU" ,    "6-BMA" ,    "9-BLA"  ,   "11-CIN" ,  "12-NAV", "13-DUL"  ,"14-PAR", "15-MEJ","24-PPE" ,"17-COC", "18-GOR",  "20-IGE" ,   "21-IGD"   ,"19-IGP" ))

#for top to bottom order
data$Ordered_SITE<-factor(data$Ordered_SITE,levels=c("19-IGP","21-IGD","20-IGE","18-GOR","17-COC","24-PPE","15-MEJ","14-PAR","13-DUL","12-NAV","11-CIN","9-BLA","6-BMA","4-NGU","3-SDB","1-SGR_SBN"))
N<-ggplot(data, aes(y=d15N, x=Ordered_SITE,fill=Region)) + 
  geom_boxplot(outlier.shape = 1,width=0.75) +theme_bw()+ guides(fill=FALSE)+
  theme(panel.grid.major.x=element_blank(),
        panel.grid.minor.x=element_blank(),
        panel.grid.major.y=element_line(colour="grey60",linetype="dashed"))+ scale_fill_manual(values=reg_palette,guide=FALSE)+xlab("Location ") +ylab("d15N") +coord_flip(ylim = c(5, 22))
N
# pdf("../Output_figures/nitrogen_boxplot.pdf", 8, 5)
# N
# dev.off()

C<-ggplot(data, aes(y=d13C, x=Ordered_SITE,fill=Region)) + 
  geom_boxplot(outlier.shape = 1,width=0.75) +theme_bw()+ guides(fill=FALSE)+
  theme(panel.grid.major.x=element_blank(),
        panel.grid.minor.x=element_blank(),
        panel.grid.major.y=element_line(colour="grey60",linetype="dashed"))+ scale_fill_manual(values=reg_palette,guide=FALSE)+xlab("Location ") +ylab("d13C")+coord_flip()
C
# pdf("../Output_figures/carbon_boxplot.pdf", 8, 5)
# C
# dev.off()
```

#### Supplemental Figure 4. See seperate script for Bayesian ellipses and convex hull areas
